### 2017-07-19 examples/profiles_examples.Rmd
Add timing examples for flights data

---
output:
  github_document:
    toc: true 
    toc_depth: 3
---

### Hypothesis tests
![h-test diagram](../figs/ht-diagram.png)

#### Examples

```{r message=FALSE, warning=FALSE}
library(okcupiddata)
library(stringr)
library(infer)
set.seed(2017)
prof_small <- profiles %>% 
  dplyr::sample_n(size = 500) %>% 
  dplyr::mutate(city = dplyr::case_when(
    str_detect(location, "san fran") ~ "san fran",
    !str_detect(location, "san fran") ~ "not san fran"
  )) %>% 
  dplyr::select(age, sex, city, drugs, height, status)
```

- `height` and `age` are numerical variables.
- `sex` has two categories (`"m"`, `"f"`)
- `city` has two categories (`"san fran"`, `"not san fran"`)
- `drugs` has three categories (`"never"`, `"sometimes"`, `"often"`) - Also has missing values
- `status` has three categories (`"single"`, `"available"`, `"seeing someone"`)

***

One numerical variable (mean)

```{r}
prof_small %>%
  specify(response = age) %>% # alt: age ~ NULL (or age ~ 1)
  hypothesize(null = "point", mu = 50) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "mean") %>% 
  visualize()
```

One numerical variable (median)

```{r}
prof_small %>%
  specify(response = age) %>% # alt: age ~ NULL (or age ~ 1)
  hypothesize(null = "point", Med = 55) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "median") %>% 
  visualize()
```

One categorical (2 level) variable

```{r}
prof_small %>%
  specify(response = sex) %>% # alt: sex ~ NULL (or sex ~ 1)
  hypothesize(null = "point", p = c("m" = .65)) %>% 
  generate(reps = 1000, type = "simulate") %>% 
  calculate(stat = "prop") %>% 
  visualize()
```

Two categorical (2 level) variables

```{r}
prof_small %>%
  specify(sex ~ city) %>% # alt: response = sex, explanatory = vs
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in props") %>% 
  visualize()
```

One categorical (>2 level) - GoF

```{r}
prof_small %>%
  specify(drugs ~ NULL) %>% # alt: response = drugs
  hypothesize(null = "point", 
              p = c("never" = .7, "sometimes" = .25, "often" = .05)) %>%
  generate(reps = 1000, type = "simulate") %>%
  calculate(stat = "Chisq") %>% 
  visualize()
```

Two categorical (>2 level) variables

```{r}
prof_small %>%
  specify(drugs ~ status) %>% # alt: response = drugs, explanatory = status
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "Chisq") %>% 
  visualize()
```

One numerical variable one categorical (2 levels) (diff in means)

```{r}
prof_small %>%
  specify(age ~ sex) %>% # alt: response = age, explanatory = sex
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means") %>% 
  visualize()
```

One numerical variable one categorical (2 levels) (diff in medians)

```{r}
prof_small %>%
  specify(age ~ sex) %>% # alt: response = age, explanatory = sex
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in medians") %>% 
  visualize()
```

One numerical one categorical (>2 levels) -  ANOVA

```{r}
prof_small %>%
  specify(age ~ status) %>% # alt: response = age, explanatory = status
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "F") %>% 
  visualize()
```

Two numerical vars - SLR

```{r}
prof_small %>%
  specify(age ~ height) %>% # alt: response = age, explanatory = height
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "slope") %>% 
  visualize()
```

### Confidence intervals

One numerical (one mean)

```{r}
prof_small %>%
  specify(response = age) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "mean") %>% 
  visualize()
```

One numerical (one median)

```{r}
prof_small %>%
  specify(response = age) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "median") %>% 
  visualize()
```

One categorical (one proportion)

```{r}
prof_small %>%
  specify(response = sex) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop", success = "f") %>% 
  visualize()
```

One numerical variable one categorical (2 levels) (diff in means)

```{r}
prof_small %>%
  specify(age ~ sex) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "diff in means") %>% 
  visualize()
```

Two categorical variables (diff in proportions)

```{r}
prof_small %>%
  specify(sex ~ city) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "diff in props") %>% 
  visualize()
```

Two numerical vars - SLR

```{r}
prof_small %>%
  specify(age ~ height) %>% 
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "slope") %>% 
  visualize()
```