### 2017-12-27 examples/two_sample_t.Rmd

---
title: "Two sample $t$ test example using `okcupiddata` `profiles` data"
author: "Chester Ismay"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: |
  %\VignetteIndexEntry{Two sample t test example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 3) 
```

## Data preparation

```{r message=FALSE, warning=FALSE}
library(okcupiddata)
library(stringr)
library(infer)
set.seed(2017)
prof_small <- profiles %>% 
  dplyr::sample_n(size = 500) %>% 
  dplyr::mutate(city = dplyr::case_when(
    str_detect(location, "san fran") ~ "san fran",
    !str_detect(location, "san fran") ~ "not san fran"
  )) %>% 
  dplyr::select(age, sex, city, drugs, height, status)
```

- `height` and `age` are numerical variables.
- `sex` has two categories (`"m"`, `"f"`)
- `city` has two categories (`"san fran"`, `"not san fran"`)
- `drugs` has three categories (`"never"`, `"sometimes"`, `"often"`) - Also has missing values
- `status` has three categories (`"single"`, `"available"`, `"seeing someone"`)

***

# One numerical variable, one categorical (2 levels)

## Calculate observed statistic

Using `t_test` in `infer`

```{r}
obs_t <- prof_small %>% 
  t_test(formula = age ~ sex) %>% 
  dplyr::select(statistic) %>% 
  dplyr::pull()
```

The observed $t$ statistic is `r obs_t`.

Or using another shortcut function in `infer`:

```{r}
obs_t <- prof_small %>% 
  t_stat(formula = age ~ sex)
```

The observed $t$ statistic is `r obs_t`.

## Randomization approach to t-statistic

```{r}
t_null_distn <- prof_small %>%
  specify(age ~ sex) %>% # alt: response = age, explanatory = sex
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "t")
t_null_distn %>% visualize(obs_stat = obs_t, direction = "two_sided")
```

## Calculate the randomization-based $p$-value

```{r}
t_null_distn %>% 
  dplyr::summarize(p_value = mean(abs(stat) >= obs_t)) %>% 
  dplyr::pull()
```

## Theoretical distribution

```{r }
prof_small %>%
  specify(age ~ sex) %>% # alt: response = age, explanatory = sex
  hypothesize(null = "independence") %>%
  # calculate(stat = "t") ## Not needed since t is implied based on variable types
  visualize(method = "theoretical", obs_stat = obs_t, direction = "two_sided")
```

## Overlay appropriate $t$ distribution on top of permuted t-statistics

```{r }
prof_small %>%
  specify(age ~ sex) %>% # alt: response = age, explanatory = sex
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "t") %>% 
  visualize(method = "both", obs_stat = obs_t, direction = "two_sided")
```

## Compute theoretical p-value

```{r}
prof_small %>% 
  t_test(formula = age ~ sex) %>% 
  dplyr::select(p_value) %>% 
  dplyr::pull()
```