### 2018-01-04 inst/doc/mtcars_examples.Rmd
fix typo in p-value and warnings.

---
title: "Examples using `mtcars` data"
author: "Chester Ismay and Andrew bray"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette
vignette: |
  %\VignetteIndexEntry{flights example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 5) 
```

## Data preparation

```{r message=FALSE, warning=FALSE}
library(nycflights13)
library(dplyr)
library(ggplot2)
library(stringr)
library(infer)
mtcars <- as.data.frame(mtcars) %>%
  mutate(cyl = factor(cyl),
         vs = factor(vs),
         am = factor(am),
         gear = factor(gear),
         carb = factor(carb))
```

***

One numerical variable (mean)

```{r}
mtcars %>%
  specify(response = mpg) %>% # formula alt: mpg ~ NULL
  hypothesize(null = "point", mu = 25) %>% 
  generate(reps = 100, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

One numerical variable (median)

```{r}
mtcars %>%
  specify(response = mpg) %>% # formula alt: mpg ~ NULL
  hypothesize(null = "point", med = 26) %>% 
  generate(reps = 100, type = "bootstrap") %>% 
  calculate(stat = "median")
```

One numerical variable (standard deviation)

```{r}
mtcars %>%
  specify(response = mpg) %>% # formula alt: mpg ~ NULL
  hypothesize(null = "point", sigma = 5) %>% 
  generate(reps = 100, type = "bootstrap") %>% 
  calculate(stat = "sd")
```

One categorical (2 level) variable

```{r}
mtcars %>%
  specify(response = am, success = "1") %>% # formula alt: am ~ NULL
  hypothesize(null = "point", p = .25) %>% 
  generate(reps = 100, type = "simulate") %>% 
  calculate(stat = "prop")
```

Two categorical (2 level) variables

```{r}
mtcars %>%
  specify(am ~ vs, success = "1") %>% # alt: response = am, explanatory = vs
  hypothesize(null = "independence") %>%
  generate(reps = 100, type = "permute") %>%
  calculate(stat = "diff in props", order = c("0", "1"))
```

One categorical (>2 level) - GoF

```{r}
mtcars %>%
  specify(cyl ~ NULL) %>% # alt: response = cyl
  hypothesize(null = "point", p = c("4" = .5, "6" = .25, "8" = .25)) %>%
  generate(reps = 100, type = "simulate") %>%
  calculate(stat = "Chisq")
```

Two categorical (>2 level) variables

```{r warning = FALSE}
mtcars %>%
  specify(cyl ~ am) %>% # alt: response = cyl, explanatory = am
  hypothesize(null = "independence") %>%
  generate(reps = 100, type = "permute") %>%
  calculate(stat = "Chisq")
```

One numerical variable one categorical (2 levels) (diff in means)

```{r}
mtcars %>%
  specify(mpg ~ am) %>% # alt: response = mpg, explanatory = am
  hypothesize(null = "independence") %>%
  generate(reps = 100, type = "permute") %>%
  calculate(stat = "diff in means", order = c("0", "1"))
```

One numerical variable one categorical (2 levels) (diff in medians)

```{r}
mtcars %>%
  specify(mpg ~ am) %>% # alt: response = mpg, explanatory = am
  hypothesize(null = "independence") %>%
  generate(reps = 100, type = "permute") %>%
  calculate(stat = "diff in medians", order = c("0", "1"))
```

One numerical one categorical (>2 levels) -  ANOVA

```{r}
mtcars %>%
  specify(mpg ~ cyl) %>% # alt: response = mpg, explanatory = cyl
  hypothesize(null = "independence") %>%
  generate(reps = 100, type = "permute") %>%
  calculate(stat = "F")
```

Two numerical vars - SLR

```{r}
mtcars %>%
  specify(mpg ~ hp) %>% # alt: response = mpg, explanatory = cyl
  hypothesize(null = "independence") %>%
  generate(reps = 100, type = "permute") %>%
  calculate(stat = "slope")
```

### Confidence intervals

One numerical (one mean)

```{r}
mtcars %>%
  specify(response = mpg) %>%
  generate(reps = 100, type = "bootstrap") %>%
  calculate(stat = "mean")
```

One numerical (one median)

```{r}
mtcars %>%
  specify(response = mpg) %>%
  generate(reps = 100, type = "bootstrap") %>%
  calculate(stat = "median")
```

One numerical (standard deviation)

```{r}
mtcars %>%
  specify(response = mpg) %>%
  generate(reps = 100, type = "bootstrap") %>%
  calculate(stat = "sd")
```

One categorical (one proportion)

```{r}
mtcars %>%
  specify(response = am, success = "1") %>%
  generate(reps = 100, type = "bootstrap") %>%
  calculate(stat = "prop")
```

One numerical variable one categorical (2 levels) (diff in means)

```{r}
mtcars %>%
  specify(mpg ~ am) %>%
  generate(reps = 100, type = "bootstrap") %>%
  calculate(stat = "diff in means", order = c("0", "1"))
```

Two categorical variables (diff in proportions)

```{r}
mtcars %>%
  specify(am ~ vs, success = "1") %>%
  generate(reps = 100, type = "bootstrap") %>%
  calculate(stat = "diff in props", order = c("0", "1"))
```

Two numerical vars - SLR

```{r}
mtcars %>%
  specify(mpg ~ hp) %>% 
  generate(reps = 100, type = "bootstrap") %>%
  calculate(stat = "slope")
```