### 2018-03-08 inst/doc/two_sample_t.Rmd

---
title: "Two sample $t$ test example using `nycflights13` `flights` data"
author: "Chester Ismay"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: |
  %\VignetteIndexEntry{Two sample t test flights example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 3) 
```

## Data preparation

```{r message=FALSE, warning=FALSE}
library(nycflights13)
library(dplyr)
library(stringr)
library(infer)
set.seed(2017)
fli_small <- flights %>% 
  sample_n(size = 500) %>% 
  mutate(half_year = case_when(
    between(month, 1, 6) ~ "h1",
    between(month, 7, 12) ~ "h2"
  )) %>% 
  mutate(day_hour = case_when(
    between(hour, 1, 12) ~ "morning",
    between(hour, 13, 24) ~ "not morning"
  )) %>% 
  select(arr_delay, dep_delay, half_year, 
         day_hour, origin, carrier)
```

* Two numeric - `arr_delay`, `dep_delay`
* Two categories 
    - `half_year` (`"h1"`, `"h2"`), 
    - `day_hour` (`"morning"`, `"not morning"`)
* Three categories - `origin` (`"EWR"`, `"JFK"`, `"LGA"`)
* Sixteen categories - `carrier`

***

# One numerical variable, one categorical (2 levels)

## Calculate observed statistic

Using `t_test` in `infer`

```{r}
obs_t <- fli_small %>% 
  t_test(formula = arr_delay ~ half_year) %>% 
  dplyr::select(statistic) %>% 
  dplyr::pull()
```

The observed $t$ statistic is `r obs_t`.

Or using another shortcut function in `infer`:

```{r}
obs_t <- fli_small %>% 
  t_stat(formula = arr_delay ~ half_year)
```

The observed $t$ statistic is `r obs_t`.

## Randomization approach to t-statistic

```{r}
t_null_distn <- fli_small %>%
  specify(arr_delay ~ half_year) %>% # alt: response = arr_delay, explanatory = half_year
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "t")
t_null_distn %>% visualize(obs_stat = obs_t, direction = "two_sided")
```

## Calculate the randomization-based $p$-value

```{r}
t_null_distn %>% 
  dplyr::summarize(p_value = mean(abs(stat) >= obs_t)) %>% 
  dplyr::pull()
```

## Theoretical distribution

```{r }
fli_small %>%
  specify(arr_delay ~ half_year) %>% # alt: response = arr_delay, explanatory = half_year
  hypothesize(null = "independence") %>%
  # calculate(stat = "t") ## Not needed since t is implied based on variable types
  visualize(method = "theoretical", obs_stat = obs_t, direction = "two_sided")
```

## Overlay appropriate $t$ distribution on top of permuted t-statistics

```{r eval=FALSE}
fli_small %>%
  specify(arr_delay ~ half_year) %>% # alt: response = arr_delay, explanatory = half_year
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "t") %>% 
  visualize(method = "both", obs_stat = obs_t, direction = "two_sided")
```

```{r echo=FALSE}
# To use same distribution calculated above
t_null_distn %>% 
  visualize(method = "both", obs_stat = obs_t, direction = "two_sided")
```

## Compute theoretical p-value

```{r}
fli_small %>% 
  t_test(formula = arr_delay ~ half_year) %>% 
  dplyr::select(p_value) %>% 
  dplyr::pull()
```