### 2018-03-30 inst/doc/chisq_test.Rmd
Add shade_color and fix two-tailed highlighting

---
title: "Chi-squared test example using `nycflights13` `flights` data"
author: "Chester Ismay"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: |
  %\VignetteIndexEntry{Chi-squared test flights example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 3) 
```

## Data preparation

```{r message=FALSE, warning=FALSE}
library(nycflights13)
library(dplyr)
library(ggplot2)
library(stringr)
library(infer)
set.seed(2017)
fli_small <- flights %>% 
  na.omit() %>% 
  sample_n(size = 500) %>% 
  mutate(season = case_when(
    month %in% c(10:12, 1:3) ~ "winter",
    month %in% c(4:9) ~ "summer"
  )) %>% 
  mutate(day_hour = case_when(
    between(hour, 1, 12) ~ "morning",
    between(hour, 13, 24) ~ "not morning"
  )) %>% 
  select(arr_delay, dep_delay, season, 
         day_hour, origin, carrier)
```

* Two numeric - `arr_delay`, `dep_delay`
* Two categories 
    - `season` (`"winter"`, `"summer"`), 
    - `day_hour` (`"morning"`, `"not morning"`)
* Three categories - `origin` (`"EWR"`, `"JFK"`, `"LGA"`)
* Sixteen categories - `carrier`

***

# One numerical variable, one categorical (2 levels)

## Calculate observed statistic

Using `chisq_test` in `infer`

```{r}
obs_chisq <- fli_small %>% 
  chisq_test(formula = origin ~ season) %>% 
  dplyr::select(statistic) %>% 
  dplyr::pull()
```

The observed $t$ statistic is `r obs_chisq`.

Or using another shortcut function in `infer`:

```{r}
obs_chisq <- fli_small %>% 
  chisq_stat(formula = origin ~ season)
```

The observed $\chi^2$ statistic is `r obs_chisq`.

## Randomization approach to $\chi^2$-statistic

```{r}
chisq_null_distn <- fli_small %>%
  specify(origin ~ season) %>% # alt: response = origin, explanatory = season
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "Chisq")
chisq_null_distn %>% visualize(obs_stat = obs_chisq, direction = "greater")
```

## Calculate the randomization-based $p$-value

```{r}
chisq_null_distn %>% 
  dplyr::summarize(p_value = mean(stat >= obs_chisq)) %>% 
  dplyr::pull()
```

## Theoretical distribution

```{r }
fli_small %>%
  specify(origin ~ season) %>% # alt: response = origin, explanatory = season
  hypothesize(null = "independence") %>%
  # calculate(stat = "Chisq") ## Not needed since Chisq is implied based on variable types
  visualize(method = "theoretical", obs_stat = obs_chisq, direction = "right")
```

## Overlay appropriate $\chi^2$ distribution on top of permuted statistics

```{r eval=FALSE}
fli_small %>%
  specify(origin ~ season) %>% # alt: response = origin, explanatory = season
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "Chisq") %>% 
  visualize(method = "both", obs_stat = obs_chisq, direction = "right")
```

```{r echo=FALSE}
# To use same distribution calculated above
chisq_null_distn %>% 
  visualize(method = "both", obs_stat = obs_chisq, direction = "right")
```

## Compute theoretical p-value

```{r}
fli_small %>% 
  chisq_test(formula = origin ~ season) %>% 
  dplyr::select(p_value) %>% 
  dplyr::pull()
```