### 2018-03-30 inst/doc/profiles_examples_both_methods.Rmd
Tweak vignettes

---
title: "Randomization and Traditional Examples using `okcupiddata` `profiles` data"
author: "Chester Ismay"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: |
  %\VignetteIndexEntry{OKCupid profiles example with resampling & theory-based methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.width = 7, fig.height = 3) 
```

# Note

There has been some recent [controversy](https://www.wired.com/2016/05/okcupid-study-reveals-perils-big-data-science/) surrounding data of this sort similar to what is given in the {okcupiddata} package. Explicit permission was obtained from OkCupid and usernames were not included in the data set for that package. Its use here is only
to check for a variety of different variable types and levels. It will remain only in this
"cutting edge" version of the package as we continue to add and test examples.

The `method = "both"` examples here do another randomization instead of using the one
with `visualize()` set to the default of `method = "randomization"`. This is to show the full
pipeline for a variety of examples, but more than likely you'll want to do something like the
following instead in practice.

```{r eval=FALSE}
null_distn <- prof_small %>%
  specify(age ~ sex) %>% # alt: response = age, explanatory = sex
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "t", order = c("m", "f"))

null_distn %>% visualize()
null_distn %>% visualize(method = "both")
```

## Data preparation

```{r message=FALSE, warning=FALSE}
library(okcupiddata)
library(stringr)
library(infer)
set.seed(2017)
prof_small <- profiles %>% 
  na.omit() %>% 
  dplyr::sample_n(size = 500) %>% 
  dplyr::mutate(city = dplyr::case_when(
    str_detect(location, "san fran") ~ "san fran",
    !str_detect(location, "san fran") ~ "not san fran"
  )) %>% 
  dplyr::select(age, sex, city, drugs, height, status)
```

- `height` and `age` are numerical variables.
- `sex` has two categories (`"m"`, `"f"`)
- `city` has two categories (`"san fran"`, `"not san fran"`)
- `drugs` has three categories (`"never"`, `"sometimes"`, `"often"`) - Also has missing values
- `status` has three categories (`"single"`, `"available"`, `"seeing someone"`)

***

## Hypothesis tests

## Implemented

### One numerical variable, one categorical (2 levels)

Randomization approach to t-statistic

```{r}
prof_small %>%
  specify(age ~ sex) %>% # alt: response = age, explanatory = sex
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "t", order = c("m", "f")) %>% 
  visualize()
```

Theoretical distribution

```{r }
prof_small %>%
  specify(age ~ sex) %>% # alt: response = age, explanatory = sex
  hypothesize(null = "independence") %>%
  # generate() is not needed since we are not doing randomization
  # calculate(stat = "t") ## Not needed since t implied based on variable types
  visualize(method = "theoretical")
```

Overlay appropriate $t$ distribution on top of permuted t-statistics

```{r }
prof_small %>%
  specify(age ~ sex) %>% # alt: response = age, explanatory = sex
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "t", order = c("m", "f")) %>% 
  visualize(method = "both")
```

### One numerical, one categorical (>2 levels) -  ANOVA

Randomization approach to F-statistic

```{r}
prof_small %>%
  specify(age ~ status) %>% # alt: response = age, explanatory = status
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "F") %>% 
  visualize()
```

Theoretical distribution

```{r}
prof_small %>%
  specify(age ~ status) %>% # alt: response = age, explanatory = status
  # generate() is not needed since we are not doing randomization
  # calculate(stat = "F") ## Not needed since F implied based on variable types
  hypothesize(null = "independence") %>%
  visualize(method = "theoretical")
```

Overlay appropriate $F$ distribution on top of permuted F-statistics

```{r}
prof_small %>%
  specify(age ~ status) %>% # alt: response = age, explanatory = status
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "F") %>% 
  visualize(method = "both")
```

## One categorical (2 level) variable

Randomization approach to z-statistic

```{r}
prof_small %>%
  specify(response = sex, success = "m") %>% 
  hypothesize(null = "point", p = c("m" = .65, "f" = .35)) %>% 
  generate(reps = 1000, type = "simulate") %>% 
  calculate(stat = "z") %>% 
  visualize()
```

Theoretical distribution

```{r}
prof_small %>%
  specify(response = sex, success = "m") %>%
  hypothesize(null = "point", p = c("m" = .65, "f" = .35)) %>% 
  # generate() is not needed since we are not doing randomization
  # calculate(stat = "z") ## Not needed since z implied based on variable types
  visualize(method = "theoretical")
```

Overlay standard normal distribution on top of permuted z-statistics

```{r}
prof_small %>%
  specify(response = sex, success = "m") %>%
  hypothesize(null = "point", p = c("m" = .65, "f" = .35)) %>% 
  generate(reps = 1000, type = "simulate") %>% 
  calculate(stat = "z") %>% 
  visualize(method = "both")
```

## Two categorical (2 level) variables

Randomization approach to z-statistic

```{r}
prof_small %>%
  specify(sex ~ city, success = "m") %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "z", order = c("san fran", "not san fran")) %>% 
  visualize()
```

Theoretical distribution

```{r}
prof_small %>%
  specify(sex ~ city, success = "m") %>%
  hypothesize(null = "independence") %>% 
  # generate() is not needed since we are not doing randomization
  # calculate(stat = "z") ## Not needed since z implied based on variable types
  visualize(method = "theoretical")
```

Overlay standard normal distribution on top of permuted z-statistics

```{r}
prof_small %>%
  specify(sex ~ city, success = "m") %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "z", order = c("san fran", "not san fran")) %>% 
  visualize(method = "both")
```

## Two categorical (>2 level) variables

Randomization approach to $\chi^2$-statistic

```{r}
prof_small %>%
  specify(drugs ~ status) %>% # alt: response = drugs, explanatory = status
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "Chisq") %>% 
  visualize()
```

Theoretical distribution

```{r}
prof_small %>%
  specify(drugs ~ status) %>% # alt: response = drugs, explanatory = status
  hypothesize(null = "independence") %>%
  # generate() is not needed since we are not doing randomization
  # calculate(stat = "Chisq") ## Not needed since Chisq implied based on variable types
  visualize(method = "theoretical")
```

Overlay standard chi-square distribution on top of permuted statistics

```{r}
prof_small %>%
  specify(drugs ~ status) %>% # alt: response = drugs, explanatory = status
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "Chisq") %>% 
  visualize(method = "both")
```

## One categorical (>2 level) - GoF

Randomization approach to $\chi^2$-statistic

```{r}
prof_small %>%
  specify(drugs ~ NULL) %>% # alt: response = drugs
  hypothesize(null = "point", 
              p = c("never" = .7, "often" = .05, "sometimes" = .25)) %>%
  generate(reps = 1000, type = "simulate") %>%
  calculate(stat = "Chisq") %>% 
  visualize()
```

Theoretical distribution

```{r}
prof_small %>%
  specify(drugs ~ NULL) %>% # alt: response = drugs
  hypothesize(null = "point", 
              p = c("never" = .7, "often" = .05, "sometimes" = .25)) %>%
  # generate() is not needed since we are not doing randomization
  # calculate(stat = "Chisq") ## Not needed since z implied based on variable types
  visualize(method = "theoretical")
```

Overlay standard chi-square distribution on top of permuted statistics

```{r}
prof_small %>%
  specify(drugs ~ NULL) %>% # alt: response = drugs
  hypothesize(null = "point", 
              p = c("never" = .7, "often" = .05, "sometimes" = .25)) %>%
  generate(reps = 1000, type = "simulate") %>%
  calculate(stat = "Chisq") %>% 
  visualize(method = "both")
```

## Two numerical vars - SLR

Randomization approach to $t$-statistic

```{r}
prof_small %>%
  specify(age ~ height) %>% # alt: response = age, explanatory = height
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "t") %>% 
  visualize()
```

Theoretical distribution

```{r}
prof_small %>%
  specify(age ~ height) %>% # alt: response = age, explanatory = height
  hypothesize(null = "independence") %>%
  # generate() is not needed since we are not doing randomization
  # calculate(stat = "t") ## Not needed since z implied based on variable types
  visualize(method = "theoretical")
```

Overlay appropriate $t$ distribution on top of permuted t-statistics

```{r}
prof_small %>%
  specify(age ~ height) %>% # alt: response = age, explanatory = height
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "t") %>% 
  visualize(method = "both")
```

# Determining whether or not to implement

Note that all code from here throughout the rest of this vignette is to think about but is not currently implemented

## One numerical variable (mean)

Randomization approach to $t$-statistic (Not implemented)

```{r eval=FALSE}
prof_small %>%
  specify(response = age) %>% # alt: age ~ NULL (or age ~ 1)
  hypothesize(null = "point", mu = 50) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "t") %>% 
  visualize()
```

Theoretical distribution (Implemented)

```{r}
prof_small %>%
  specify(response = age) %>% # alt: age ~ NULL (or age ~ 1)
  hypothesize(null = "point", mu = 50) %>% 
#  generate(reps = 1000, type = "bootstrap") %>% 
#  calculate(stat = "t") %>% 
  visualize(method = "theoretical")
```

Overlay appropriate $t$ distribution on top of permuted t-statistics (Not implemented)

```{r eval=FALSE}
prof_small %>%
  specify(response = age) %>% # alt: age ~ NULL (or age ~ 1)
  hypothesize(null = "point", mu = 50) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "t") %>% 
  visualize(method = "both")
```

## Confidence intervals

### To Do (one mean $t$ stat)
### To Do (Also overlay appropriate t distn)

One numerical (one mean)

```{r eval=FALSE}
prof_small %>%
  specify(response = age) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "mean") %>% 
  visualize()
```

### To Do (one prop $z$ stat)
### To Do (Also overlay std normal distn)

One categorical (one proportion)

```{r eval=FALSE}
prof_small %>%
  specify(response = sex) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop", success = "f") %>% 
  visualize()
```

### To Do (Also overlay appropriate t distn)
### To Do (Shift t stats to be centered at 0 for bootstrapping)

One numerical variable one categorical (2 levels) (diff in means)

```{r eval=FALSE}
prof_small %>%
  specify(age ~ sex) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "t") %>% 
  visualize()
```

### To Do (two prop $z$ stat)
### To Do (Also overlay appropriate std normal distn)
### To Do (Shift z stats to be centered at 0 for bootstrapping)

Two categorical variables (diff in proportions)

```{r eval=FALSE}
prof_small %>%
  specify(sex ~ city) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "diff in props") %>% 
  visualize()
```

### To Do (slope $t$ stat)
### To Do (Also overlay appropriate t distn)
### To Do (Shift t stats to be centered at 0 for bootstrapping)

Two numerical vars - SLR

```{r eval=FALSE}
prof_small %>%
  specify(age ~ height) %>% 
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "slope") %>% 
  visualize()
```