### 2018-05-13 inst/doc/profiles_examples.Rmd

---
title: "Randomization Examples using `okcupiddata` `profiles` data"
author: "Chester Ismay"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: |
  %\VignetteIndexEntry{OKCupid profiles example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 5) 
```

# Note

There has been some recent [controversy](https://www.wired.com/2016/05/okcupid-study-reveals-perils-big-data-science/) 
surrounding data of this sort similar to what is given in the {okcupiddata}
package. Explicit permission was obtained from OkCupid and usernames were not
included in the data set for that package. Its use here is only
to check for a variety of different variable types and levels. It will remain 
only in this "cutting edge" version of the package as we continue to add 
and test examples.

## Data preparation

```{r message=FALSE, warning=FALSE}
library(okcupiddata)
#remotes::install_github("andrewpbray/infer", ref = "develop")
library(infer)

set.seed(2017)
prof_small <- profiles %>% 
  na.omit() %>% 
  dplyr::sample_n(size = 500) %>% 
  dplyr::mutate(city = dplyr::case_when(
    stringr::str_detect(location, "san fran") ~ "san fran",
    !stringr::str_detect(location, "san fran") ~ "not san fran"
  )) %>% 
  dplyr::select(age, sex, city, drugs, height, status)
```

- `height` and `age` are numerical variables.
- `sex` has two categories (`"m"`, `"f"`)
- `city` has two categories (`"san fran"`, `"not san fran"`)
- `drugs` has three categories (`"never"`, `"sometimes"`, `"often"`) - Also has missing values
- `status` has three categories (`"single"`, `"available"`, `"seeing someone"`)

***

## Hypothesis tests

One numerical variable (mean)

```{r}
x_bar <- prof_small %>% 
  dplyr::summarize(mean(age)) %>% 
  dplyr::pull()
prof_small %>%
  specify(response = age) %>% # alt: age ~ NULL (or age ~ 1)
  hypothesize(null = "point", mu = 34) %>% 
  generate(reps = 1000) %>% 
  calculate(stat = "mean") %>% 
  visualize(obs_stat = x_bar, direction = "both")
```

One numerical variable (median)

```{r}
x_tilde <- prof_small %>% 
  dplyr::summarize(median(age)) %>% 
  dplyr::pull()
prof_small %>%
  specify(response = age) %>% # alt: age ~ NULL
  hypothesize(null = "point", med = 29) %>% 
  generate(reps = 1000) %>% 
  calculate(stat = "median") %>% 
  visualize(obs_stat = x_tilde, direction = "both")
```

One categorical (2 level) variable

```{r}
p_hat <- prof_small %>% 
  dplyr::summarize(mean(sex == "f")) %>% 
  dplyr::pull()
prof_small %>%
  specify(response = sex, success = "f") %>%
  hypothesize(null = "point", p = c("f" = .35)) %>% 
  generate(reps = 1000) %>% 
  calculate(stat = "prop") %>% 
  visualize(obs_stat = p_hat, direction = "right")
```

Two categorical (2 level) variables

```{r}
obs_diff <- prof_small %>% 
  dplyr::group_by(city) %>% 
  dplyr::summarize(prop = mean(sex == "m")) %>%
  dplyr::summarize(diff(prop)) %>%
  dplyr::pull()
prof_small %>%
  specify(sex ~ city, success = "m") %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000) %>%
  calculate(stat = "diff in props", order = c("san fran", "not san fran")) %>% 
  visualize(obs_stat = obs_diff, direction = "both")
```

One categorical (>2 level) - GoF

```{r}
Chisq_hat <- chisq.test(table(prof_small$drugs))$stat
prof_small %>%
  specify(drugs ~ NULL) %>% # alt: response = drugs
  hypothesize(null = "point", 
              p = c("never" = .7, "sometimes" = .25, "often" = .05)) %>%
  generate(reps = 1000) %>%
  calculate(stat = "Chisq") %>% 
  # Only "right" or "greater" as options here
  visualize(obs_stat = Chisq_hat, direction = "right")
```

Two categorical (>2 level) variables

```{r}
Chisq_hat <- prof_small %>% chisq_stat(formula = drugs ~ status)
prof_small %>%
  specify(drugs ~ status) %>% # alt: response = drugs, explanatory = status
  hypothesize(null = "independence") %>%
  generate(reps = 1000) %>%
  calculate(stat = "Chisq") %>% 
  # Only "right" or "greater" as options here
  visualize(obs_stat = Chisq_hat, direction = "right")
```

One numerical variable, one categorical (2 levels) (diff in means)

```{r}
obs_diff <- prof_small %>% 
  dplyr::group_by(sex) %>% 
  dplyr::summarize(mean_age = mean(age)) %>% 
  dplyr::summarize(diff(mean_age)) %>% 
  dplyr::pull()
prof_small %>%
  specify(age ~ sex) %>% # alt: response = age, explanatory = sex
  hypothesize(null = "independence") %>%
  generate(reps = 1000) %>%
  calculate(stat = "diff in means", order = c("m", "f")) %>% 
  visualize(obs_stat = obs_diff, direction = "both")
```

One numerical variable, one categorical (2 levels) (diff in medians)

```{r}
obs_diff <- prof_small %>% 
  dplyr::group_by(sex) %>% 
  dplyr::summarize(median_age = median(age)) %>% 
  dplyr::summarize(diff(median_age)) %>% 
  dplyr::pull()
prof_small %>%
  specify(age ~ sex) %>% # alt: response = age, explanatory = sex
  hypothesize(null = "independence") %>%
  generate(reps = 1000) %>%
  calculate(stat = "diff in medians", order = c("m", "f")) %>% 
  visualize(obs_stat = obs_diff, direction = "two_sided")
```

One numerical one categorical (>2 levels) -  ANOVA

```{r}
F_hat <- anova(
               aov(formula = age ~ status, data = prof_small)
               )$`F value`[1]
prof_small %>%
  specify(age ~ status) %>% # alt: response = age, explanatory = status
  hypothesize(null = "independence") %>%
  generate(reps = 1000) %>%
  calculate(stat = "F") %>% 
  # Only "right" or "greater" as options here
  visualize(obs_stat = F_hat, direction = "greater")
```

Two numerical vars - SLR

```{r}
slope_hat <- lm(age ~ height, data = prof_small) %>% 
  broom::tidy() %>% 
  dplyr::filter(term == "height") %>% 
  dplyr::select(estimate) %>% 
  dplyr::pull()
prof_small %>%
  specify(age ~ height) %>% # alt: response = age, explanatory = height
  hypothesize(null = "independence") %>%
  generate(reps = 1000) %>%
  calculate(stat = "slope") %>% 
  visualize(obs_stat = slope_hat, direction = "both")
```

## Confidence intervals

### Confidence interval shading is not implemented yet

One numerical (one mean)

```{r}
prof_small %>%
  specify(response = age) %>%
  generate(reps = 1000) %>%
  calculate(stat = "mean") %>% 
  visualize()
```

One numerical (one median)

```{r}
prof_small %>%
  specify(response = age) %>%
  generate(reps = 1000) %>%
  calculate(stat = "median") %>% 
  visualize()
```

One categorical (one proportion)

```{r}
prof_small %>%
  specify(response = sex, success = "f") %>%
  generate(reps = 1000) %>%
  calculate(stat = "prop") %>% 
  visualize()
```

One numerical variable one categorical (2 levels) (diff in means)

```{r}
prof_small %>%
  specify(age ~ sex) %>%
  generate(reps = 1000) %>%
  calculate(stat = "diff in means", order = c("m", "f")) %>% 
  visualize()
```

Two categorical variables (diff in proportions)

```{r}
prof_small %>%
  specify(sex ~ city, success = "m") %>%
  generate(reps = 1000) %>%
  calculate(stat = "diff in props", order = c("san fran", "not san fran")) %>% 
  visualize()
```

Two numerical vars - SLR

```{r}
prof_small %>%
  specify(age ~ height) %>% 
  generate(reps = 1000) %>%
  calculate(stat = "slope") %>% 
  visualize()
```