### 2020-10-12 README.Rmd
adjust figure path
The visualize output in the README has a bad path on the websites.

---
output: github_document
---

# infer R Package <img src="https://github.com/tidymodels/infer/blob/master/figs/infer_gnome.png?raw=true" align="right" width=250 />

------------------------------------------------------------------------

<!--figs/infer.svg-->
<!--http://www.r-pkg.org/badges/version/infer-->
<!--figs/master.svg-->
<!--https://img.shields.io/codecov/c/github/tidymodels/infer/master.svg-->

[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/infer)](https://cran.r-project.org/package=infer) [![Travis-CI Build Status](https://travis-ci.org/tidymodels/infer.svg?branch=master)](https://travis-ci.org/tidymodels/infer) [![Coverage Status](https://img.shields.io/codecov/c/github/tidymodels/infer/master.svg)](https://codecov.io/github/tidymodels/infer/?branch=master)

The objective of this package is to perform statistical inference using an expressive statistical grammar that coheres with the `tidyverse` design framework. The package is centered around 4 main verbs, supplemented with many utilities to visualize and extract value from their outputs.

+ `specify()` allows you to specify the variable, or relationship between variables, that you're interested in.
+ `hypothesize()` allows you to declare the null hypothesis.
+ `generate()` allows you to generate data reflecting the null hypothesis.
+ `calculate()` allows you to calculate a distribution of statistics from the generated data to form the null distribution.

To learn more about the principles underlying the package design, see `vignette("infer")`.

```{r load-package, echo = FALSE, message = FALSE, warning = FALSE}
library(devtools)
devtools::load_all()
```

```{r diagram, echo=FALSE}
knitr::include_graphics("https://raw.githubusercontent.com/tidymodels/infer/master/figs/ht-diagram.png")
```

### Installation

------------------------------------------------------------------------

To install the current stable version of `infer` from CRAN:

```{r, eval = FALSE}
install.packages("infer")
```

To install the developmental stable version of `infer`, make sure to install `remotes` first. The `pkgdown` website for the package is at <https://infer.tidymodels.org>.

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("tidymodels/infer")
```

### Contributing

------------------------------------------------------------------------

We welcome others helping us make this package as user-friendly and efficient as possible. Please review our [contributing](https://github.com/tidymodels/infer/blob/master/CONTRIBUTING.md) and [conduct](https://github.com/tidymodels/infer/blob/master/CONDUCT.md) guidelines. By participating in this project you agree to abide by its terms.

For questions and discussions about tidymodels packages, modeling, and machine learning, please [post on RStudio Community](https://rstd.io/tidymodels-community). If you think you have encountered a bug, please [submit an issue](https://github.com/tidymodels/infer/issues). Either way, learn how to create and share a [reprex](https://rstd.io/reprex) (a minimal, reproducible example), to clearly communicate about your code. Check out further details on [contributing guidelines for tidymodels packages](https://www.tidymodels.org/contribute/) and [how to get help](https://www.tidymodels.org/help/).

### Examples

------------------------------------------------------------------------

These examples are pulled from the "Full infer Pipeline Examples" vignette, accessible by calling `vignette("observed_stat_examples")`. They make use of the `gss` dataset supplied by the package, providing a sample of data from the [General Social Survey](https://gss.norc.org). The data looks like this:

```{r load-gss, warning = FALSE, message = FALSE}
# load in the dataset
data(gss)

# take a glimpse at it
str(gss)
```

As an example, we'll run an analysis of variance on `age` and `partyid`, testing whether the age of a respondent is independent of their political party affiliation.

Calculating the observed statistic,

```{r, message = FALSE, warning = FALSE}
F_hat <- gss %>% 
  specify(age ~ partyid) %>%
  calculate(stat = "F")
```

Then, generating the null distribution,

```{r, message = FALSE, warning = FALSE}
null_distn <- gss %>%
   specify(age ~ partyid) %>%
   hypothesize(null = "independence") %>%
   generate(reps = 1000, type = "permute") %>%
   calculate(stat = "F")
```

Visualizing the observed statistic alongside the null distribution,

```{r viz, message = FALSE, warning = FALSE}
visualize(null_distn) +
  shade_p_value(obs_stat = F_hat, direction = "greater")
```

Calculating the p-value from the null distribution and observed statistic,

```{r, message = FALSE, warning = FALSE}
null_distn %>%
  get_p_value(obs_stat = F_hat, direction = "greater")
```

Note that the formula and non-formula interfaces  (i.e. `age ~ partyid` vs. `response = age, explanatory =  partyid`) work for all implemented inference procedures in `infer`. Use whatever is more natural for you. If you will be doing modeling using functions like `lm()` and `glm()`, though, we recommend you begin to use the formula `y ~ x` notation as soon as possible.

Other resources are available in the package vignettes! See `vignette("observed_stat_examples")` for more examples like the one above, and `vignette("infer")` for discussion of the underlying principles of the package design.