### 2021-06-14 README.Rmd
add alt text to README figures

---
output: github_document
---

# infer R Package <img src="https://github.com/tidymodels/infer/blob/master/figs/infer_gnome.png?raw=true" align="right" width=280 />

<!--figs/infer.svg-->
<!--http://www.r-pkg.org/badges/version/infer-->
<!--figs/master.svg-->
<!--https://img.shields.io/codecov/c/github/tidymodels/infer/master.svg-->

[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/infer)](https://cran.r-project.org/package=infer)[![Coverage Status](https://img.shields.io/codecov/c/github/tidymodels/infer/master.svg)](https://codecov.io/github/tidymodels/infer/?branch=master)

The objective of this package is to perform statistical inference using an expressive statistical grammar that coheres with the `tidyverse` design framework. The package is centered around 4 main verbs, supplemented with many utilities to visualize and extract value from their outputs.

+ `specify()` allows you to specify the variable, or relationship between variables, that you're interested in.
+ `hypothesize()` allows you to declare the null hypothesis.
+ `generate()` allows you to generate data reflecting the null hypothesis.
+ `calculate()` allows you to calculate a distribution of statistics from the generated data to form the null distribution.

To learn more about the principles underlying the package design, see `vignette("infer")`.

```{r load-package, echo = FALSE, message = FALSE, warning = FALSE}
library(devtools)
devtools::load_all()
```

```{r diagram, echo = FALSE, fig.cap = " ", fig.alt = "A diagram showing four steps to carry out randomization-based inference: specify hypothesis, generate data, calculate statistic, and visualize. From left to right, each step is connected by an arrow, while the diagram indicates that generating data and calculating statistics can happen iteratively."}
knitr::include_graphics("https://raw.githubusercontent.com/tidymodels/infer/master/figs/ht-diagram.png")
```

### Installation

------------------------------------------------------------------------

To install the current stable version of `infer` from CRAN:

```{r, eval = FALSE}
install.packages("infer")
```

To install the developmental stable version of `infer`, make sure to install `remotes` first. The `pkgdown` website for this version is at [infer.tidymodels.org](https://infer.tidymodels.org/).

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("tidymodels/infer")
```

### Contributing

------------------------------------------------------------------------

We welcome others helping us make this package as user-friendly and efficient as possible. Please review our [contributing](https://github.com/tidymodels/infer/blob/master/CONTRIBUTING.md) and [conduct](https://github.com/tidymodels/infer/blob/master/CONDUCT.md) guidelines. By participating in this project you agree to abide by its terms.

For questions and discussions about tidymodels packages, modeling, and machine learning, please [post on RStudio Community](https://rstd.io/tidymodels-community). If you think you have encountered a bug, please [submit an issue](https://github.com/tidymodels/infer/issues). Either way, learn how to create and share a [reprex](https://rstd.io/reprex) (a minimal, reproducible example), to clearly communicate about your code. Check out further details on [contributing guidelines for tidymodels packages](https://www.tidymodels.org/contribute/) and [how to get help](https://www.tidymodels.org/help/).

### Examples

------------------------------------------------------------------------

These examples are pulled from the "Full infer Pipeline Examples" vignette, accessible by calling `vignette("observed_stat_examples")`. They make use of the `gss` dataset supplied by the package, providing a sample of data from the [General Social Survey](https://gss.norc.org). The data looks like this:

```{r load-gss, warning = FALSE, message = FALSE}
# load in the dataset
data(gss)

# take a glimpse at it
str(gss)
```

As an example, we'll run an analysis of variance on `age` and `partyid`, testing whether the age of a respondent is independent of their political party affiliation.

Calculating the observed statistic,

```{r, message = FALSE, warning = FALSE}
F_hat <- gss %>% 
  specify(age ~ partyid) %>%
  calculate(stat = "F")
```

Then, generating the null distribution,

```{r, message = FALSE, warning = FALSE}
null_distn <- gss %>%
   specify(age ~ partyid) %>%
   hypothesize(null = "independence") %>%
   generate(reps = 1000, type = "permute") %>%
   calculate(stat = "F")
```

Visualizing the observed statistic alongside the null distribution,

```{r viz, message = FALSE, warning = FALSE, eval = FALSE}
visualize(null_distn) +
  shade_p_value(obs_stat = F_hat, direction = "greater")
```

```{r viz-graphic, message = FALSE, warning = FALSE, echo = FALSE, fig.cap = " ", fig.alt = "A histogram showing a distribution of F statistics, right-tailed and centered around one half. The x axis ranges from zero to five. The region of the histogram to the right of the observed statistic, just above two, is shaded red to represent the p-value."}
knitr::include_graphics("https://raw.githubusercontent.com/tidymodels/infer/master/README_files/figure-gfm/viz-1.png")
```

Calculating the p-value from the null distribution and observed statistic,

```{r, message = FALSE, warning = FALSE}
null_distn %>%
  get_p_value(obs_stat = F_hat, direction = "greater")
```

Note that the formula and non-formula interfaces  (i.e. `age ~ partyid` vs. `response = age, explanatory =  partyid`) work for all implemented inference procedures in `infer`. Use whatever is more natural for you. If you will be doing modeling using functions like `lm()` and `glm()`, though, we recommend you begin to use the formula `y ~ x` notation as soon as possible.

Other resources are available in the package vignettes! See `vignette("observed_stat_examples")` for more examples like the one above, and `vignette("infer")` for discussion of the underlying principles of the package design.