### 2025-12-12 man-roxygen/seeds.Rmd

# Reproducibility

When using the infer package for research, or in other cases when exact reproducibility is a priority, be sure the set the seed for R's random number generator. infer will respect the random seed specified in the `set.seed()` function, returning the same result when `generate()`ing data given an identical seed. For instance, we can calculate the difference in mean `age` by `college` degree status using the `gss` dataset from 10 versions of the `gss` resampled with permutation using the following code.

```{r, include = FALSE}
library(infer)
```

```{r}
set.seed(1)

gss |>
  specify(age ~ college) |>
  hypothesize(null = "independence") |>
  generate(reps = 5, type = "permute") |>
  calculate("diff in means", order = c("degree", "no degree"))
```

Setting the seed to the same value again and rerunning the same code will produce the same result.

```{r}
# set the seed
set.seed(1)

gss |>
  specify(age ~ college) |>
  hypothesize(null = "independence") |>
  generate(reps = 5, type = "permute") |>
  calculate("diff in means", order = c("degree", "no degree"))
```

Please keep this in mind when writing infer code that utilizes resampling with `generate()`.